**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-07 Schloss_PacBio16S_PeerJ_2016.Rmd
Update unique/precluster error analysis code

```{r precluster}
summarizeError <- function(folder, method){
  file1 <- paste("analysis/", folder, "/", folder, ".mock1.", method, ".error.summary", sep="")
  file2 <- paste("analysis/", folder, "/", folder, ".mock2.", method, ".error.summary", sep="")
  file3 <- paste("analysis/", folder, "/", folder, ".mock3.", method, ".error.summary", sep="")

  summary <- rbind(read.table(file=file1, header=T, row.names=1), read.table(file=file2, header=T, row.names=1), read.table(file=file3, header=T, row.names=1))
  nochim <- summary[summary$numparents==1,]
  error <- sum(nochim$weight * nochim$mismatches) / sum(nochim$weight * nochim$total)
  return(error)
}

unique.error <- unlist(lapply(regions, summarizeError, "unique"))
precluster.error <- unlist(lapply(regions, summarizeError, "precluster"))
final.error <- cbind(unique.error, precluster.error)
rownames(final.error) <- regions

#    unique.error precluster.error
#v4   0.002277391      0.001418871
#v13  0.009081866      0.007722886
#v35  0.009719159      0.008257968
#v15  0.004740243      0.002645747
#v16  0.007130228      0.005676505
#v19  0.005158478      0.003163881

reduction <- 100*(1-final.error[,"precluster.error"]/final.error[,"unique.error"])
rnd.reduction <- round(reduction, 0)
min.rnd.red <- min(rnd.reduction)
max.rnd.red <- max(rnd.reduction)

min.region <- paste(pretty.region[names(rnd.reduction[rnd.reduction==min.rnd.red])],collapse=" and ")

max.region <- pretty.region[[names(rnd.reduction[rnd.reduction==max.rnd.red])]]
```

***Pre-clustering sequences to further reduce sequencing noise.***
Previously, we implemented a pre-clustering algorithm where sequences were sorted
by their abundance in decreasing order and rare sequences are clustered with a
more abundant sequence if the rare sequences have fewer mismatches than a
defined threshold when compared to the more abundant sequence. The recommended
threshold was a 1-nt difference per 100-nt of sequence data. For example, the
threshold for 250 bp fragment from the V4 region would be 2 nt or 14 for the
1458 bp V1-V9 fragments. This approach removes residual PCR and sequencing
errors while not overwhelming the resolution needed to identify OTUs that are
based on a 3% distance threshold. The tradeoff of this approach is that one would
unable to differentiate V1-V9 sequences that truly differed by less than 14 nt.
When we applied
this approach to our PacBio data, we observed a reduction in the error rate
between `r min.rnd.red` (`r min.region`) and `r max.rnd.red`% (`r max.region`).
The final error rates varied between
`r round(min(100*final.error[,"precluster.error"]), digits=2)`
(`r pretty.region[names(which.min(final.error[,"precluster.error"]))]`)
and `r round(max(100*final.error[,"precluster.error"]), digits=2)`%
(`r pretty.region[names(which.max(final.error[,"precluster.error"]))]`); the
full-length, V1-V9, fragments had an error rate of
`r round(100* final.error["v19","precluster.error"], digits=2)`%
(Figure 2B). These error rates are 7-40
times higher than what we have previously observed using the 454 and
Illumina MiSeq platforms (0.02%){Schloss, 2011 #2466; Kozich, 2013 #2719}

>>>>

```{r precluster}
get_error_rate <- function(file_name){
	error_summary <- read.table(file=file_name, header=T, stringsAsFactors=FALSE)
	nochim <- error_summary[error_summary$numparents==1,]
    sum(nochim$weight * nochim$mismatches) / sum(nochim$weight * nochim$total)
}

regions <- c("V4", "V1V3", "V1V5", "V1V6", "V1V9", "V3V5")

unique_file_names <- paste0("data/mothur_pool/", regions, ".mock.screen.unique.good.filter.unique.error.summary")

unique_error <- sapply(unique_file_names, get_error_rate)

precluster_file_names <- paste0("data/mothur_pool/", regions,".mock.screen.unique.good.filter.unique.precluster.error.summary")

precluster_error <- sapply(precluster_file_names, get_error_rate)

pooled_error <- cbind(unique = unique_error, precluster = precluster_error)
rownames(pooled_error) <- regions

#          unique   precluster
#V4   0.003205151 0.0015805612
#V1V3 0.002529819 0.0006841053
#V1V5 0.001987322 0.0001938975
#V1V6 0.001910158 0.0003086235
#V1V9 0.001944654 0.0002743820
#V3V5 0.002076498 0.0003532300

reduction <- 100*(1-pooled_error[,"precluster"]/pooled_error[,"unique"])

min.region <- paste(pretty.region[names(rnd.reduction[rnd.reduction==min.rnd.red])],collapse=" and ")
max.region <- pretty.region[[names(rnd.reduction[rnd.reduction==max.rnd.red])]]
```

***Pre-clustering sequences to further reduce sequencing noise.***
Previously, we implemented a pre-clustering algorithm where sequences were sorted
by their abundance in decreasing order and rare sequences are clustered with a
more abundant sequence if the rare sequences have fewer mismatches than a
defined threshold when compared to the more abundant sequence [refs]. The recommended
threshold was a 1-nt difference per 100-nt of sequence data. For example, the
threshold for 250 bp fragment from the V4 region would be 2 nt or 14 for the
1458 bp V1-V9 fragments. This approach removes residual PCR and sequencing
errors while not overwhelming the resolution needed to identify OTUs that are
based on a 3% distance threshold. The tradeoff of this approach is that one would
unable to differentiate V1-V9 sequences that truly differed by less than 14 nt.
When we applied this approach to our PacBio data, we observed a reduction in the error rate
between `r min.rnd.red` (`r min.region`) and `r max.rnd.red`% (`r max.region`).
The final error rates varied between
`r round(100*min(pooled_error[,"precluster"]), digits=2)` (`r names(which.min(pooled_error[,"precluster"]))`) and `r round(100*max(pooled_error[,"precluster"]), digits=2)`% (`r names(which.max(pooled_error[,"precluster"]))`).
 The full-length (i.e. V1V9) fragments had an error rate of
`r round(100* pooled_error["V1V9","precluster"], digits=2)`%
(Figure 2B), this is similar to what we have previously observed using the 454 and
Illumina MiSeq platforms (0.02%){Schloss, 2011 #2466; Kozich, 2013 #2719}
